
---
title: "PRS dendrogram clusters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(reticulate)
#use_python("/home/oma21/miniconda3/envs/enm/bin/python")
#use_condaenv('enm')
#setwd('../')
library(RColorBrewer)
```

# Read data
Read data in python, from pickle file saved earlier.
```{python}
import matplotlib.pyplot as plt
import matplotlib as mpl
import networkx as nx
import numpy as np
import pandas as pd
import os
from enm.Enm import Enm
from enm.utils import *
import scipy.cluster.hierarchy as sch
with open(f'data/interim/pcc.pickle','rb') as f:
    e_pcc = pickle.load(f)
pcc_df = e_pcc.df
pcc_gc = e_pcc.graph_gc
pcc_gc_pos = (dict(pcc_gc.nodes(data='pos')))


```

Use scipy distance calculation and pass the square form of the distance matrix to R.
This is done because R's `dist` function is significantly slower than scipy.

```{python}
prs_mat = e_pcc.prs_mat
quantile_threshold =  0.95
method = 'ward'
metric = 'seuclidean'
q99 = np.quantile(prs_mat, quantile_threshold)
prs_mat_cl = copy.deepcopy(prs_mat)
prs_mat_cl[prs_mat_cl > q99] = q99
prs_mat_cl_orig = prs_mat_cl
row_dist_orig = sch.distance.pdist((prs_mat),metric = metric)
col_dist_orig = sch.distance.pdist((prs_mat.T),metric = metric)
row_dist_orig_sqr = sch.distance.squareform(row_dist_orig)
col_dist_orig_sqr = sch.distance.squareform(col_dist_orig)

row_dist = sch.distance.pdist(prs_mat_cl,metric = metric)
col_dist = sch.distance.pdist(prs_mat_cl.T,metric = metric)

row_dist_sqr = sch.distance.squareform(row_dist)
col_dist_sqr = sch.distance.squareform(col_dist)

```
# Plot heatmap
Using the clustering with `seuclidian` distance and `ward` linkage, plot the heatmap with row and column colors are the GO term annotations found from effector and sensor analysis

```{r}
library(tidyverse)
library(ComplexHeatmap)
cluster_method = 'ward.D2'
dist_row = as.dist(py$row_dist_sqr)
hclust_row = hclust(dist_row, cluster_method)
dist_col = as.dist(py$col_dist_sqr)
hclust_col = hclust(dist_col, cluster_method)
row_dend = as.dendrogram(hclust_row)%>% dendextend::color_branches(k=2, col = c('black', '#f15a24'))
col_dend = as.dendrogram(hclust_col)%>%  dendextend::color_branches(k=2, col = c('#f7931e', 'black'))
#row_dend <- row_dend %>%

# go_color_map = unlist(append(py$effector_colors_dict,py$sensor_colors_dict))
# go_color_map['none'] = 'white'
# clrs = go_color_map# map(py$eff_sens_go_list, ~ifelse(is.null(.x), 'none',.x))%>%unlist()

# rowColors = rowAnnotation(rowColors=py$e_pcc$df%>%left_join(py$effector_pcc)%>%select(go_term)%>%pull%>%map(~ifelse(is.na(.x), 'none',.x))%>%unlist(),
#                           col=list(rowColors = go_color_map ),
#                           #width=unit(100,'mm'),
#                           simple_anno_size=unit(3,'cm'),
#                           show_annotation_name=FALSE,
#                           show_legend=FALSE)
# colColors = columnAnnotation(colColors=py$e_pcc$df%>%left_join(py$sensors_pcc)%>%select(go_term)%>%pull%>%map(~ifelse(is.null(.x), 'none',
#                                                                                                                       ifelse(is.nan(.x),'none',.x)))%>%unlist(),
#                           col=list(colColors = go_color_map ),
#                           simple_anno_size=unit(3,'cm'),
#                           #annotation_height=unit(20,'mm'),
#                           show_annotation_name=FALSE,
#                           show_legend=FALSE)
rowAnnot = HeatmapAnnotation(Effectiveness=anno_lines(rowSums(py$prs_mat), border=FALSE,
                            width = unit(0.2, 'in'),
                            axis=FALSE),
                            which='row',
                            #name='', 
                            gp = gpar(col='blue'),
                            annotation_name_side = 'top',
                            annotation_name_rot= 90,
                            border=FALSE,
                            #annotation_name_gp= gpar(fontsize = 0)
                            show_annotation_name=FALSE,
                          show_legend=FALSE
)
colAnnot = HeatmapAnnotation(Sensitivity=anno_lines(colSums(py$prs_mat), 
                            axis = FALSE, border=FALSE,height = unit(0.2, 'in')), 
                             which='col', 
                            # name='',
                            gp = gpar(col='blue'),
                            #annotation_name_gp= gpar(fontsize = 0)
                            annotation_name_side = 'left',
                            annotation_name_rot=0,
                            border=FALSE,
                            show_annotation_name=FALSE,
                          show_legend=FALSE
)

png('tmp_heatmap.png', width = 3, height=2.5, units='in', res=150 )
hm<-Heatmap(py$prs_mat_cl, 
        cluster_rows = row_dend,#as.dendrogram(hclust_row),
        cluster_columns= col_dend,#as.dendrogram(hclust_col),
        col=RColorBrewer::brewer.pal(9,'YlGnBu'),
        row_dend_width = unit(0.2,'in'),
        column_dend_height = unit(0.2,'in'),
        right_annotation=rowAnnot,
#        left_annotation = rowColors,
        bottom_annotation=colAnnot,
#        top_annotation=colColors,
show_heatmap_legend = FALSE,
)
draw(hm ,padding = unit(c(0, 0.5, 0, 0), "in"))

dev.off()

png('tmp_heatmap_lgn.png', width = 3, height=2.5, units='in', res=150 )

lgd = Legend(title = "", 
            at = c(0, 1e-4,2e-4,3e-4,4e-4), col_fun = circlize::colorRamp2(c(0, 1e-4,2e-4,3e-4,4e-4), RColorBrewer::brewer.pal(5,'YlGnBu')) , labels_gp = gpar(font=6), title_position = 'leftcenter-rot', legend_width=unit(0.05, 'in'),labels_rot=45)
#
draw(lgd, x = unit(0.1, "npc"), y = unit(0.8, "npc"))

dev.off()
```

